---
title: "EPA workflow"
author: "Anaïs Ladoy"
date: "5/21/2018"
output: 
  html_document:
    toc: true # table of content true
    toc_depth: 3  # upto three depths of headings (specified by #, ## and ###)
    theme: united
---

The present notebook is dedicated to the workflow implementation for the EPA's air quality data. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, results=FALSE, warning=FALSE}

library(tidyverse) # (ggplot2, dplyr, readr, etc)

library(sf) # Simple Features

library(USAboundaries) # Boundaries for geographical units in the United States of America (U.S. Census Bureau)

```


# Get the data

Download the dataset from the available sources providing EPA's data.

  * Historical hourly data (run once)
  * Hourly data from the previous day (will become an automatic process)
  * Metadata about the monitoring stations


```{r, results=FALSE, warning=FALSE}

#ftp://ftp.airnowapi.org/
#files.airnowtech.org

#HOURLY DATA FOR THE PREVIOUS DAY
date <- Sys.Date()-1
date <- str_replace_all(date,pattern='-',replacement = '')

hour <- c("00","01","02","03","04","05","06","07","08","09","10","11","12","13","14","15","16","17","18","19","20","21","22","23")
fieldnames.hourly <- c("vdate","vtime","AQSID","site.name","GMT.offset","parameter.name","report.units",
                "value","data.source")
data <- data.frame()

for(i in 1:24)
{
hourly_dat <- read.delim(paste0("https://s3-us-west-1.amazonaws.com//files.airnowtech.org/airnow/",substring(date,0,4),"/",date,"/HourlyData_",date,hour[i],".dat"), sep="|", header=FALSE, col.names = fieldnames.hourly)
data <- rbind(data,hourly_dat)
}

#Backup
write_csv(data,paste0("Data/AirNow/HourlyValues/",date,".csv"))

#DOWNLOAD MONITORING SITES METADATA FOR THE PREVIOUS DAY (up-to-date)

fieldnames.meta <- c("AQSID","parameter.name","site.code","site.name","status","agency.id","agency.name","EPA.region","latitude","longitude","elevation","GMT.offset","country.code","CMSA.code","CMSA.name","MSA.code","MSA.name","state.code","state.name","county.code","county.name","city.code","city.name")

MS <- read.delim(paste0("https://s3-us-west-1.amazonaws.com//files.airnowtech.org/airnow/",substring(date,0,4),"/",date,"/monitoring_site_locations.dat"), sep="|", header=FALSE, col.names = fieldnames.meta)


#HISTORICAL DATA (RUN ONCE)

#Ask Luc for the Hourly data and not daily data

```


```{r, results=FALSE, warning=FALSE}
#aqs.epa.org
#For now only criteria but see the list of parameters group



# for(i in 1:24)
# {
# hourly_dat <- read.delim(paste0("https://s3-us-west-1.amazonaws.com//files.airnowtech.org/airnow/",substring(date,0,4),"/",date,"/HourlyData_",date,hour[i],".dat"), sep="|", header=FALSE, col.names = fieldnames.hourly)
# data <- rbind(data,hourly_dat)
# }
# r<-read_csv("https://aqs.epa.gov/api/rawData?user=anais.ladoy@epfl.ch&pw=saffrongoose78&format=AQCSV&pc=CRITERIA&param=&bdate=20180401&edate=20180401&state=17&county=31&dur=1")


bdate <- "20180301"
edate <- "20180401"

#Function to extract data for a specific time period and for a specific county.
aqs_dat <- function(bdate,edate,state,county) {
  return (read_csv(paste0("https://aqs.epa.gov/api/rawData?user=anais.ladoy@epfl.ch&pw=saffrongoose78&format=AQCSV&pc=CRITERIA&param=&bdate=",bdate,"&edate=",edate,"&state=",state,"&county=",county,"&dur=1")))
}

data <- data.frame()

#IL counties
counties.IL <- c('031','097','043','197','091','111','089','093','063')
for (i in 1:length(counties.IL)) {
  r.IL <- aqs_dat(bdate,edate,"17",counties.IL[i])
  data <- rbind(data,r.IL) 
}

#IN counties ('LAKE','PORTER','NEWTON','JASPER','LA PORTE','STARKE','PULASKI')
counties.IN <- c('089','127','111','073','091','149','131')
for (i in 1:length(counties.IN)) {
  r.IN <- aqs_dat(bdate,edate,"92",counties.IN[i])
  data <- rbind(data,r.IN) 
}

#WI counties ('RACINE','WALWORTH','KENOSHA','MILWAUKEE','WAUKESHA')
counties.WI <- c('101','127','059','079','133')
for (i in 1:length(counties.WI)) {
  r.WI <- aqs_dat(bdate,edate,"55",counties.WI[i])
  data <- rbind(data,r.WI) 
}

data.bk <- data
data <- data.bk

#Remove empty values (NA lines), must be equal to 21 (nb of requests=nb of counties)
if(nrow(data %>% filter(is.na(site) | site=='END OF FILE')) == 21){
  data <- data %>% filter(!is.na(site) & site!='END OF FILE')
}
{
  print('There are more NA lines than expected. You may check again.')
}


#Time


#Send to timescale
fieldnames.AQS <- c('site_id','dat_status','act_code','datetime','param_code','dur_code','freq','value','unit','qc','poc','lat','lon','datum','elev','meth_code','mpc','mpc_value','uncert','qualif')

```




# Daily Dump

First connection to the spatial database, the raw file (csv) is pushed in order to have a copy of the dataset.

  1. Filter the monitoring stations to keep the ones surrounding Chicago (15 neighboring counties distributed through Wisconsin, Indiana and Illinois)
  2. In the metadata dataframe, add informations (manually) about the spatial scale and monitoring objective of each site using the different state's EPA reports
  3. Check if the metadata has changed since the last dump and if it's the case, overwrite the metadata file in the DB with the new file
  4. Add the hourly data file in the DB (raw values filtered with the Monitoring Stations)
  

```{r, results=FALSE, warning=FALSE}

#FILTER THE MONITORING STATIONS
MS <- MS %>% filter((state.name=='IL' & county.name %in% c('COOK','LAKE','DU PAGE','WILL','KANKAKEE','MC HENRY','KANE','KENDALL','GRUNDY')) | (state.name=="WI" & county.name %in% c('RACINE','WALWORTH','KENOSHA','MILWAUKEE','WAUKESHA')) | (state.name=="IN" & county.name %in% c('LAKE','PORTER','NEWTON','JASPER','LA PORTE','STARKE','PULASKI') ))  %>% droplevels() %>% st_as_sf(coords=c('longitude','latitude'),agr = 'identity',crs = 4326) %>% st_transform(crs=32616)

#Filter the data to keep only the MS in the study area, drop unused levels (from the previous datagrame) and convert to a tibble dataframe (optimal data format)
data <- data %>% filter(AQSID %in% MS$AQSID) %>% droplevels() %>% as_tibble()

#ADD INFORMATION IN METADATA
#From official reports (Illinois, Indiana, Wisconsin)

head(data)

#Check if the metadata has changed ?
#st_transform(counties, crs=32616)

```


```{r, results=FALSE, warning=FALSE}
#MS where we don't have any data
print(MS %>% filter(AQSID %in% setdiff(levels(MS$AQSID), levels(data$AQSID))))
```

Here is the list of the monitoring stations registered in the metadata file but for which we don't have any data. Except for two stations (LaPorte-E. Lincolnway, CHI-SP), it corresponds to stations that are no longer active.

No insights about the reason why we don't have any data for these two active stations :
* 17-031-0057 : Springfield Pump Station (1745 N. Springfield Ave.)
* 18-091-0010 : LaPorte - E. Lincolnway (2011 E. Lincolnway)

# Data Wrangling

Rearrange the dataframe in order to have an optimal format (tidy) for the DB integration.

  1. Check for any duplicates or other "dirty data" problems
  2. Remove unecessary variables to keep only ID/timestamp/parameter/value (to be consistent with the AoT format)
  3. Reshape the dataframe in a tidy format
  4. Complete the time serie to make missing values explicit (insert a NA row for missing timestamp)

don't forget to convert timestamp

```{r, results=FALSE, warning=FALSE}

#Map the counties
limits.counties <- us_counties() %>% filter((state_name=='Illinois' & name %in% c('Cook','Lake','DuPage','Will','Kankakee','McHenry','Kane','Kendall','Grundy')) | (state_name=="Wisconsin" & name %in% c('Racine','Walworth','Kenosha','Milwaukee','Waukesha')) | (state_name=="Indiana" & name %in% c('Lake','Porter','Newton','Jasper','LaPorte','Starke','Pulaski'))) %>% st_transform(counties, crs=32616) 

ggplot() + geom_sf(data=limits.counties,mapping=aes(fill = state_name)) + geom_sf(data=MS) + facet_wrap(~parameter.name, nrow=3) + ggtitle('Monitoring Sites location according to the AirNow metadata file (both active and inactive stations)')


d <- data[duplicated(data),]
```



# Before Modeling

Tasks that need to be implemented once the data will be on the dataset and before any modeling.

  1. Outliers / Anomalies detection
  2. Imputation of missing values


